---
title: "week4_Introduction to tidyverse"
author: "Macarena López-Mayorga"
date: 2026-03-27
format: html
---
First, we need to indicate where is located the file

```{r}
#| eval: FALSE
FilePath <- file.path("C:/Users/Macarena/CytometryInR/course/week04/04_IntroToTidyverse/data/Dataset.csv")
FilePath
```

```{r}
#| eval: FALSE
library(utils)
```


```{r}
#| eval: FALSE
Data <- read.csv(file=FilePath, check.names=FALSE)
colnames(Data)
```
 [1] "bid"               "timepoint"         "Condition"         "Date"              "infant_sex"        "ptype"            
 [7] "root"              "singletsFSC"       "singletsSSC"       "singletsSSCB"      "CD45"              "NotMonocytes"     
[13] "nonDebris"         "lymphocytes"       "live"              "Dump+"             "Dump-"             "Tcells"           
[19] "Vd2+"              "Vd2-"              "Va7.2+"            "Va7.2-"            "CD4+"              "CD4-"             
[25] "CD8+"              "CD8-"              "Tcells_count"      "lymphocytes_count" "Monocytes"         "Debris"           
[31] "CD45_count"

when we put check.names into TRUE, then the "+" or "-" symbols dissapear

```{r}
#| eval: FALSE
Data <- read.csv(file=FilePath, check.names=TRUE)
colnames(Data)
```
 [1] "bid"               "timepoint"         "Condition"         "Date"              "infant_sex"        "ptype"            
 [7] "root"              "singletsFSC"       "singletsSSC"       "singletsSSCB"      "CD45"              "NotMonocytes"     
[13] "nonDebris"         "lymphocytes"       "live"              "Dump."             "Dump..1"           "Tcells"           
[19] "Vd2."              "Vd2..1"            "Va7.2."            "Va7.2..1"          "CD4."              "CD4..1"           
[25] "CD8."              "CD8..1"            "Tcells_count"      "lymphocytes_count" "Monocytes"         "Debris"           
[31] "CD45_count"   

To see some rows:
```{r}
#| eval: FALSE
head(Data, 3)
```
bid timepoint Condition       Date infant_sex  ptype    root singletsFSC singletsSSC singletsSSCB      CD45 NotMonocytes nonDebris
1 INF0052         0      Ctrl 2025-07-26       Male HEU-hi 2098368     1894070     1666179      1537396 0.5952943    0.8820349 0.8627649
2 INF0100         0      Ctrl 2025-07-26       Male HEU-lo 2020184     1791890     1697083      1579098 0.9106762    0.9052256 0.8602660
3 INF0100         4      Ctrl 2025-07-26       Male HEU-lo 1155040     1033320      875465       845446 0.9705765    0.9845400 0.9578793
  lymphocytes      live      Dump+     Dump-    Tcells        Vd2+      Vd2-     Va7.2+    Va7.2-      CD4+      CD4-      CD8+
1   0.6420138 0.9020581 0.21090996 0.6911482 0.2804264 0.008120361 0.9918796 0.01448070 0.9773989 0.6341164 0.3432825 0.2734826
2   0.2145848 0.8908981 0.06252775 0.8283703 0.6748298 0.007265620 0.9927344 0.01577499 0.9769594 0.6119112 0.3650482 0.3357696
3   0.7403110 0.8757665 0.20023803 0.6755285 0.6119129 0.004651313 0.9953487 0.01579402 0.9795547 0.6639621 0.3155925 0.2862104
        CD8- Tcells_count lymphocytes_count  Monocytes     Debris CD45_count
1 0.06979990       164771            587573 0.11796509 0.13723513     915203
2 0.02927858       208241            308583 0.09477437 0.13973396    1438047
3 0.02938209       371723            607477 0.01545999 0.04212072     820570

To visualize the table, we can click into the grid, in the rigth panel, or using the next code:
```{r}
#| eval: FALSE
View(Data)
```
To see the type of data:

```{r}
#| eval: FALSE
class(Data)
```
[1] "data.frame"

And see the variables:
```{r}
#| eval: FALSE
str(Data)
```
'data.frame':	196 obs. of  31 variables:
 $ bid              : chr  "INF0052" "INF0100" "INF0100" "INF0100" ...
 $ timepoint        : int  0 0 4 9 0 4 9 4 9 0 ...
 $ Condition        : chr  "Ctrl" "Ctrl" "Ctrl" "Ctrl" ...
 $ Date             : chr  "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26" ...
 $ infant_sex       : chr  "Male" "Male" "Male" "Male" ...
 $ ptype            : chr  "HEU-hi" "HEU-lo" "HEU-lo" "HEU-lo" ...
 $ root             : int  2098368 2020184 1155040 358624 1362216 1044808 1434840 972056 1521928 2363512 ...
 $ singletsFSC      : int  1894070 1791890 1033320 328624 1206309 917398 1265022 875707 1359574 2136616 ...
 $ singletsSSC      : int  1666179 1697083 875465 289327 1032946 735579 988445 767323 1175755 1875394 ...
 $ singletsSSCB     : int  1537396 1579098 845446 276289 982736 685592 940454 718000 1097478 1732620 ...
 $ CD45             : num  0.595 0.911 0.971 0.982 0.957 ...
 $ NotMonocytes     : num  0.882 0.905 0.985 0.986 0.956 ...
 $ nonDebris        : num  0.863 0.86 0.958 0.941 0.841 ...
 $ lymphocytes      : num  0.642 0.215 0.74 0.651 0.705 ...
 $ live             : num  0.902 0.891 0.876 0.915 0.895 ...
 $ Dump+            : num  0.2109 0.0625 0.2002 0.2147 0.3383 ...
 $ Dump-            : num  0.691 0.828 0.676 0.701 0.557 ...
 $ Tcells           : num  0.28 0.675 0.612 0.631 0.44 ...
 $ Vd2+             : num  0.00812 0.00727 0.00465 0.01135 0.00475 ...
 $ Vd2-             : num  0.992 0.993 0.995 0.989 0.995 ...
 $ Va7.2+           : num  0.0145 0.0158 0.0158 0.017 0.0133 ...
 $ Va7.2-           : num  0.977 0.977 0.98 0.972 0.982 ...
 $ CD4+             : num  0.634 0.612 0.664 0.438 0.739 ...
 $ CD4-             : num  0.343 0.365 0.316 0.534 0.243 ...
 $ CD8+             : num  0.273 0.336 0.286 0.486 0.195 ...
 $ CD8-             : num  0.0698 0.0293 0.0294 0.0476 0.0476 ...
 $ Tcells_count     : int  164771 208241 371723 111552 291777 271870 487937 220634 415867 184930 ...
 $ lymphocytes_count: int  587573 308583 607477 176662 663667 510730 726238 451047 710964 652155 ...
 $ Monocytes        : num  0.118 0.0948 0.0155 0.0145 0.0444 ...
 $ Debris           : num  0.1372 0.1397 0.0421 0.0587 0.1592 ...
 $ CD45_count       : int  915203 1438047 820570 271304 940733 675857 921660 701657 1066884 1017713 ...

To see that using glimpse function (if we don't know from what package is, we can use the next code to figure out)
 ```{r}
 #| eval: FALSE
 ??glimpse
 ```

 It's from dplyr package, so we have to activate it
 
 ```{r}
 #| eval: FALSE
 library(dplyr)
glimpse(Data)
 ```
 Rows: 196
Columns: 31
$ bid               <chr> "INF0052", "INF0100", "INF0100", "INF0100", "INF0179", "INF0179", "INF0179", "INF0186", "INF0186", "INF0052", …
$ timepoint         <int> 0, 0, 4, 9, 0, 4, 9, 4, 9, 0, 0, 4, 9, 0, 4, 9, 4, 9, 0, 0, 4, 0, 4, 9, 4, 9, 0, 4, 9, 0, 9, 0, 4, 9, 0, 4, 9,…
$ Condition         <chr> "Ctrl", "Ctrl", "Ctrl", "Ctrl", "Ctrl", "Ctrl", "Ctrl", "Ctrl", "Ctrl", "PPD", "PPD", "PPD", "PPD", "PPD", "PP…
$ Date              <chr> "2025-07-26", "2025-07-26", "2025-07-26", "2025-07-26", "2025-07-26", "2025-07-26", "2025-07-26", "2025-07-26"…
$ infant_sex        <chr> "Male", "Male", "Male", "Male", "Male", "Male", "Male", "Female", "Female", "Male", "Male", "Male", "Male", "M…
$ ptype             <chr> "HEU-hi", "HEU-lo", "HEU-lo", "HEU-lo", "HU", "HU", "HU", "HEU-hi", "HEU-hi", "HEU-hi", "HEU-lo", "HEU-lo", "H…
$ root              <int> 2098368, 2020184, 1155040, 358624, 1362216, 1044808, 1434840, 972056, 1521928, 2363512, 2049112, 1063496, 7883…
$ singletsFSC       <int> 1894070, 1791890, 1033320, 328624, 1206309, 917398, 1265022, 875707, 1359574, 2136616, 1821676, 946587, 714198…
$ singletsSSC       <int> 1666179, 1697083, 875465, 289327, 1032946, 735579, 988445, 767323, 1175755, 1875394, 1717636, 796056, 626387, …
$ singletsSSCB      <int> 1537396, 1579098, 845446, 276289, 982736, 685592, 940454, 718000, 1097478, 1732620, 1597085, 767297, 600011, 1…
$ CD45              <dbl> 0.5952943, 0.9106762, 0.9705765, 0.9819573, 0.9572591, 0.9858006, 0.9800160, 0.9772382, 0.9721234, 0.5873838, …
$ NotMonocytes      <dbl> 0.8820349, 0.9052256, 0.9845400, 0.9855070, 0.9556272, 0.9622894, 0.9805210, 0.9490663, 0.9581698, 0.8619837, …
$ nonDebris         <dbl> 0.8627649, 0.8602660, 0.9578793, 0.9412615, 0.8407837, 0.9499243, 0.9657759, 0.9189704, 0.9186763, 0.8429685, …
$ lymphocytes       <dbl> 0.6420138, 0.2145848, 0.7403110, 0.6511588, 0.7054786, 0.7556776, 0.7879674, 0.6428312, 0.6663930, 0.6408044, …
$ live              <dbl> 0.9020581, 0.8908981, 0.8757665, 0.9153242, 0.8952140, 0.9673056, 0.9297420, 0.8756515, 0.8749529, 0.9009254, …
$ `Dump+`           <dbl> 0.21090996, 0.06252775, 0.20023803, 0.21469246, 0.33831877, 0.36947311, 0.23863114, 0.32516789, 0.22875701, 0.…
$ `Dump-`           <dbl> 0.6911482, 0.8283703, 0.6755285, 0.7006317, 0.5568953, 0.5978325, 0.6911109, 0.5504837, 0.6461959, 0.6934931, …
$ Tcells            <dbl> 0.2804264, 0.6748298, 0.6119129, 0.6314431, 0.4396437, 0.5323165, 0.6718693, 0.4891597, 0.5849340, 0.2835676, …
$ `Vd2+`            <dbl> 0.008120361, 0.007265620, 0.004651313, 0.011348967, 0.004753630, 0.005160555, 0.002631487, 0.009300470, 0.0179…
$ `Vd2-`            <dbl> 0.9918796, 0.9927344, 0.9953487, 0.9886510, 0.9952464, 0.9948394, 0.9973685, 0.9906995, 0.9820592, 0.9925918, …
$ `Va7.2+`          <dbl> 0.014480704, 0.015774991, 0.015794019, 0.017023451, 0.013321818, 0.012461838, 0.013563226, 0.038770090, 0.0367…
$ `Va7.2-`          <dbl> 0.9773989, 0.9769594, 0.9795547, 0.9716276, 0.9819246, 0.9823776, 0.9838053, 0.9519294, 0.9452878, 0.9775212, …
$ `CD4+`            <dbl> 0.6341164, 0.6119112, 0.6639621, 0.4378944, 0.7392563, 0.5876264, 0.6994141, 0.5724730, 0.5119209, 0.6340345, …
$ `CD4-`            <dbl> 0.3432825, 0.3650482, 0.3155925, 0.5337331, 0.2426682, 0.3947512, 0.2843912, 0.3794565, 0.4333669, 0.3434867, …
$ `CD8+`            <dbl> 0.2734826, 0.3357696, 0.2862104, 0.4861231, 0.1950634, 0.3571376, 0.2580784, 0.3555526, 0.4138799, 0.2744119, …
$ `CD8-`            <dbl> 0.06979990, 0.02927858, 0.02938209, 0.04761008, 0.04760485, 0.03761357, 0.02631282, 0.02390384, 0.01948700, 0.…
$ Tcells_count      <int> 164771, 208241, 371723, 111552, 291777, 271870, 487937, 220634, 415867, 184930, 211987, 326378, 238021, 294549…
$ lymphocytes_count <int> 587573, 308583, 607477, 176662, 663667, 510730, 726238, 451047, 710964, 652155, 314717, 544883, 366784, 663169…
$ Monocytes         <dbl> 0.11796509, 0.09477437, 0.01545999, 0.01449297, 0.04437285, 0.03771064, 0.01947898, 0.05093372, 0.04183023, 0.…
$ Debris            <dbl> 0.13723513, 0.13973396, 0.04212072, 0.05873854, 0.15921627, 0.05007568, 0.03422412, 0.08102962, 0.08132374, 0.…
$ CD45_count        <int> 915203, 1438047, 820570, 271304, 940733, 675857, 921660, 701657, 1066884, 1017713, 1447451, 745037, 589379

This function is similar to str() function but more tidy and with the dimensions of the table at the top. To see only the dimensions, we can also use de dim() function, which maintains the row followed by column position convention of base R (first row, second column)

```{r}
#| eval: FALSE
dim(Data)
```
[1] 196  31

To verify the column value type (character, for instance), we can use the functions is.character, is.numeric or is.logical:

```{r}
#| eval: FALSE
is.character(Data$bid)
```
[1] TRUE

```{r}
#| eval: FALSE
is.numeric(Data$bid)
```
[1] FALSE

```{r}
#| eval: FALSE
is.character(Data$Tcells_count)
```
[1] FALSE
For numeric columns using the is.numeric() function, we can also be subtype specific using either is.integer() or is.double():

```{r}
#| eval: FALSE
is.numeric(Data$Tcells_count)
```
[1] TRUE

```{r}
#| eval: FALSE
is.integer(Data$Tcells_count)
```
[1] TRUE


```{r}
#| eval: FALSE
is.double(Data$Tcells_count)
```
[1] FALSE

Like with keywords, special characters (as & or spaces) will need tick marks to run the function

```{r}
#| eval: FALSE
is.numeric(Data$`CD8-`)
```
[1] TRUE


```{r}
#| eval: FALSE
library(dplyr)
```
The select() function is used to “select” a column from a data.frame type object. The simplest way to use it is giving the name of the data.frame object in first place and then, the name of the column, surrounded by " ":

```{r}
#| eval: FALSE
DateColumn <- select(Data, "Date")
DateColumn[1:10,]
```
[1] "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26"

This results in the column being selected, resulting in the new object containing only that subsetted out column from the original Data object.

In general, the function select() is used with the implementation of pipe operators, not like we used previously:

```{r}
#| eval: FALSE
DateColumn <- Data |> select("Date")
DateColumn[1:10,]
```
 [1] "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26" "2025-07-26"

 The main advantage of the pipe operators is that we can link different functions. For example, the nrow() function determine the number of rows. We can link the function select() and nrow() with the pipe operators:
 
 ```{r}
 #| eval: FALSE
 Data |> select("Date") |> nrow()
 ```
 [1] 196

 CURIOUS NOTE: Despite we need "" for install.packages() function, for instance, in select() function, the function still run without the "" in date:

 ```{r}
 #| eval: FALSE
 Data |> select(Date) |> head(3)
 ```
         Date
1 2025-07-26
2 2025-07-26
3 2025-07-26

This is due to dplyr; It's now in charge and it's able to infer that Date is a column, with no need of "".
For select more than one column and position them anywhere within the data.frame:

```{r}
#| eval: FALSE
Subset <- Data |> select(bid, timepoint, Condition, Tcells, `CD8+`, `CD4+`)
head(Subset, 3)
```
      bid timepoint Condition    Tcells      CD8+      CD4+
1 INF0052         0      Ctrl 0.2804264 0.2734826 0.6341164
2 INF0100         0      Ctrl 0.6748298 0.3357696 0.6119112
3 INF0100         4      Ctrl 0.6119129 0.2862104 0.6639621

The order in which we selected the columns will dictate their position in the subsetted data.frame object:

```{r}
#| eval: FALSE
Subset <- Data |> select(bid, Tcells, `CD8+`, `CD4+`, timepoint, Condition)
head(Subset, 3)
```
      bid    Tcells      CD8+      CD4+ timepoint Condition
1 INF0052 0.2804264 0.2734826 0.6341164         0      Ctrl
2 INF0100 0.6748298 0.3357696 0.6119112         0      Ctrl
3 INF0100 0.6119129 0.2862104 0.6639621         4      Ctrl

The relocate() function is used when we want to move just one column. We can do this with select() function too, but it would be a waste of time. Let’s say we wanted to move the Tcells column from its current location to the second column position (right after the bid column). The line of code to do so would look like:

```{r}
#| eval: FALSE
Data |> relocate(Tcells, .after=bid) |> head(3)
```
      bid    Tcells timepoint Condition       Date infant_sex  ptype    root singletsFSC singletsSSC singletsSSCB      CD45 NotMonocytes
1 INF0052 0.2804264         0      Ctrl 2025-07-26       Male HEU-hi 2098368     1894070     1666179      1537396 0.5952943    0.8820349
2 INF0100 0.6748298         0      Ctrl 2025-07-26       Male HEU-lo 2020184     1791890     1697083      1579098 0.9106762    0.9052256
3 INF0100 0.6119129         4      Ctrl 2025-07-26       Male HEU-lo 1155040     1033320      875465       845446 0.9705765    0.9845400
  nonDebris lymphocytes      live      Dump+     Dump-        Vd2+      Vd2-     Va7.2+    Va7.2-      CD4+      CD4-      CD8+
1 0.8627649   0.6420138 0.9020581 0.21090996 0.6911482 0.008120361 0.9918796 0.01448070 0.9773989 0.6341164 0.3432825 0.2734826
2 0.8602660   0.2145848 0.8908981 0.06252775 0.8283703 0.007265620 0.9927344 0.01577499 0.9769594 0.6119112 0.3650482 0.3357696
3 0.9578793   0.7403110 0.8757665 0.20023803 0.6755285 0.004651313 0.9953487 0.01579402 0.9795547 0.6639621 0.3155925 0.2862104
        CD8- Tcells_count lymphocytes_count  Monocytes     Debris CD45_count
1 0.06979990       164771            587573 0.11796509 0.13723513     915203
2 0.02927858       208241            308583 0.09477437 0.13973396    1438047
3 0.02938209       371723            607477 0.01545999 0.04212072     820570

Similar to what we saw with select(), this approach can also be used for more than 1 column:

```{r}
#| eval: FALSE
Data |> relocate(Tcells, Monocytes, .after=bid) |> head(3)
```
      bid    Tcells  Monocytes timepoint Condition       Date infant_sex  ptype    root singletsFSC singletsSSC singletsSSCB      CD45
1 INF0052 0.2804264 0.11796509         0      Ctrl 2025-07-26       Male HEU-hi 2098368     1894070     1666179      1537396 0.5952943
2 INF0100 0.6748298 0.09477437         0      Ctrl 2025-07-26       Male HEU-lo 2020184     1791890     1697083      1579098 0.9106762
3 INF0100 0.6119129 0.01545999         4      Ctrl 2025-07-26       Male HEU-lo 1155040     1033320      875465       845446 0.9705765
  NotMonocytes nonDebris lymphocytes      live      Dump+     Dump-        Vd2+      Vd2-     Va7.2+    Va7.2-      CD4+      CD4-
1    0.8820349 0.8627649   0.6420138 0.9020581 0.21090996 0.6911482 0.008120361 0.9918796 0.01448070 0.9773989 0.6341164 0.3432825
2    0.9052256 0.8602660   0.2145848 0.8908981 0.06252775 0.8283703 0.007265620 0.9927344 0.01577499 0.9769594 0.6119112 0.3650482
3    0.9845400 0.9578793   0.7403110 0.8757665 0.20023803 0.6755285 0.004651313 0.9953487 0.01579402 0.9795547 0.6639621 0.3155925
       CD8+       CD8- Tcells_count lymphocytes_count     Debris CD45_count
1 0.2734826 0.06979990       164771            587573 0.13723513     915203
2 0.3357696 0.02927858       208241            308583 0.13973396    1438047
3 0.2862104 0.02938209       371723            607477 0.04212072     820570
