---
Title: "Gating sets_Homeworks"
Author: "Macarena López-Mayorga"
date: 2026-04-30
format: html
---
# Problem 1
First of all, we are going to indicate the file path of the fcs files:

```{r}
Folder <- file.path("C:/Users/Macarena/CytometryInR/course/05_GatingSets/data")
fcs_files <- list.files(Folder, pattern=".fcs", full.names=TRUE)
fcs_files
```
[1] "C:/Users/Macarena/CytometryInR/course/05_GatingSets/data/2025_07_26_AB_02_INF052_00_Ctrl.fcs"
[2] "C:/Users/Macarena/CytometryInR/course/05_GatingSets/data/2025_07_26_AB_02_INF052_00_SEB.fcs" 
[3] "C:/Users/Macarena/CytometryInR/course/05_GatingSets/data/2025_07_26_AB_02_INF100_00_Ctrl.fcs"
[4] "C:/Users/Macarena/CytometryInR/course/05_GatingSets/data/2025_07_26_AB_02_INF100_00_SEB.fcs" 
[5] "C:/Users/Macarena/CytometryInR/course/05_GatingSets/data/2025_07_26_AB_02_INF179_00_Ctrl.fcs"
[6] "C:/Users/Macarena/CytometryInR/course/05_GatingSets/data/2025_07_26_AB_02_INF179_00_SEB.fcs" 

I need to activate the packages that I am going to use:
```{r}
library(flowCore)
library(lobstr)
library(flowWorkspace)
```

To read the 6 files:
```{r}
flowSet <- read.flowSet(files=fcs_files, truncate_max_range=FALSE, transformation=FALSE)
flowSet
```

To stablish the GatingSet
```{r}
GatingSet1 <- GatingSet(flowSet)
GatingSet1
```

Then, I need to call CytoML:
```{r}
library(CytoML)
```

  Then, we need to identify the FlowJo files:
```{r}
Folder <- file.path("C:/Users/Macarena/CytometryInR/course/05_GatingSets/data")
FlowJoWsp <- list.files(path=Folder, pattern=".wsp", full=TRUE)
FlowJoWsp
```

To select the file "opened":
```{r}
OpenedWsp <- FlowJoWsp[stringr::str_detect(FlowJoWsp, "Opened")]
OpenedWsp
```
To set up the intermediate object:
```{r}
ws <- open_flowjo_xml(OpenedWsp)
ws
```
To read from the FlowJo file to a GatingSet:
```{r}
GS <- flowjo_to_gatingset(ws=ws, name=1, path=Folder, additional.keys="GROUPNAME")
GS
```

We need to activate bench package:
```{r}
library(bench)
```

To retrieve counts of cells within the individual gates:
```{r}
Data <- gs_pop_get_count_fast(GS)
head(Data, 5)
```
                 name                         Population                        Parent Count ParentCount
               <char>                             <char>                        <char> <int>       <int>
1: 00_Ctrl.fcs_INF052                           /Scatter                          root 10000       10000
2: 00_Ctrl.fcs_INF052                  /Scatter/Singlets                      /Scatter  9986       10000
3: 00_Ctrl.fcs_INF052             /Scatter/Singlets/Live             /Scatter/Singlets  9708        9986
4: 00_Ctrl.fcs_INF052      /Scatter/Singlets/Live/Tcells        /Scatter/Singlets/Live  9708        9708
5: 00_Ctrl.fcs_INF052 /Scatter/Singlets/Live/Tcells/CD4+ /Scatter/Singlets/Live/Tcells  6151        9708

We need to activate dplyr to work in the Tidyverse:
```{r}
library(dplyr)
```

I include the percentage from the parent gate:
```{r}
TidyData <- Data |> mutate(Percent_Parent_Gate=(Count/ParentCount)*100)
TidyData
```
                   name                                    Population                             Parent Count ParentCount
                 <char>                                        <char>                             <char> <int>       <int>
  1: 00_Ctrl.fcs_INF052                                      /Scatter                               root 10000       10000
  2: 00_Ctrl.fcs_INF052                             /Scatter/Singlets                           /Scatter  9986       10000
  3: 00_Ctrl.fcs_INF052                        /Scatter/Singlets/Live                  /Scatter/Singlets  9708        9986
  4: 00_Ctrl.fcs_INF052                 /Scatter/Singlets/Live/Tcells             /Scatter/Singlets/Live  9708        9708
  5: 00_Ctrl.fcs_INF052            /Scatter/Singlets/Live/Tcells/CD4+      /Scatter/Singlets/Live/Tcells  6151        9708
 ---                                                                                                                      
110:  00_SEB.fcs_INF179            /Scatter/Singlets/Live/Tcells/CD8+      /Scatter/Singlets/Live/Tcells  1871        9402
111:  00_SEB.fcs_INF179 /Scatter/Singlets/Live/Tcells/CD8+/TNFa+IFNg+ /Scatter/Singlets/Live/Tcells/CD8+     0        1871
112:  00_SEB.fcs_INF179 /Scatter/Singlets/Live/Tcells/CD8+/TNFa+IFNg- /Scatter/Singlets/Live/Tcells/CD8+     7        1871
113:  00_SEB.fcs_INF179 /Scatter/Singlets/Live/Tcells/CD8+/TNFa-IFNg+ /Scatter/Singlets/Live/Tcells/CD8+    20        1871
114:  00_SEB.fcs_INF179 /Scatter/Singlets/Live/Tcells/CD8+/TNFa-IFNg- /Scatter/Singlets/Live/Tcells/CD8+  1844        1871
     Percent_Parent_Gate
                   <num>
  1:         100.0000000
  2:          99.8600000
  3:          97.2161025
  4:         100.0000000
  5:          63.3601154
 ---                    
110:          19.9000213
111:           0.0000000
112:           0.3741315
113:           1.0689471
114:          98.5569214

I decided to rounded the values obtained:
```{r}
TidyData <- TidyData |> mutate(Percent_Parent_Gate_rounded=round(Percent_Parent_Gate, 2))
head(TidyData, 5)
```
                  name                                    Population                             Parent Count ParentCount
                 <char>                                        <char>                             <char> <int>       <int>
  1: 00_Ctrl.fcs_INF052                                      /Scatter                               root 10000       10000
  2: 00_Ctrl.fcs_INF052                             /Scatter/Singlets                           /Scatter  9986       10000
  3: 00_Ctrl.fcs_INF052                        /Scatter/Singlets/Live                  /Scatter/Singlets  9708        9986
  4: 00_Ctrl.fcs_INF052                 /Scatter/Singlets/Live/Tcells             /Scatter/Singlets/Live  9708        9708
  5: 00_Ctrl.fcs_INF052            /Scatter/Singlets/Live/Tcells/CD4+      /Scatter/Singlets/Live/Tcells  6151        9708
 ---                                                                                                                      
110:  00_SEB.fcs_INF179            /Scatter/Singlets/Live/Tcells/CD8+      /Scatter/Singlets/Live/Tcells  1871        9402
111:  00_SEB.fcs_INF179 /Scatter/Singlets/Live/Tcells/CD8+/TNFa+IFNg+ /Scatter/Singlets/Live/Tcells/CD8+     0        1871
112:  00_SEB.fcs_INF179 /Scatter/Singlets/Live/Tcells/CD8+/TNFa+IFNg- /Scatter/Singlets/Live/Tcells/CD8+     7        1871
113:  00_SEB.fcs_INF179 /Scatter/Singlets/Live/Tcells/CD8+/TNFa-IFNg+ /Scatter/Singlets/Live/Tcells/CD8+    20        1871
114:  00_SEB.fcs_INF179 /Scatter/Singlets/Live/Tcells/CD8+/TNFa-IFNg- /Scatter/Singlets/Live/Tcells/CD8+  1844        1871
     Percent_Parent_Gate
                   <num>
  1:         100.0000000
  2:          99.8600000
  3:          97.2161025
  4:         100.0000000
  5:          63.3601154
 ---                    
110:          19.9000213
111:           0.0000000
112:           0.3741315
113:           1.0689471
114:          98.5569214
> TidyData <- TidyData |> mutate(Percent_Parent_Gate_rounded=round(Percent_Parent_Gate, 2))
+ head(TidyData, 5)
                 name                         Population                        Parent Count ParentCount Percent_Parent_Gate
               <char>                             <char>                        <char> <int>       <int>               <num>
1: 00_Ctrl.fcs_INF052                           /Scatter                          root 10000       10000           100.00000
2: 00_Ctrl.fcs_INF052                  /Scatter/Singlets                      /Scatter  9986       10000            99.86000
3: 00_Ctrl.fcs_INF052             /Scatter/Singlets/Live             /Scatter/Singlets  9708        9986            97.21610
4: 00_Ctrl.fcs_INF052      /Scatter/Singlets/Live/Tcells        /Scatter/Singlets/Live  9708        9708           100.00000
5: 00_Ctrl.fcs_INF052 /Scatter/Singlets/Live/Tcells/CD4+ /Scatter/Singlets/Live/Tcells  6151        9708            63.36012
   Percent_Parent_Gate_rounded
                         <num>
1:                      100.00
2:                       99.86
3:                       97.22
4:                      100.00
5:                       63.36

And now, I removed the previous column no rounded:
```{r}
TidyData <- TidyData |> select (!Percent_Parent_Gate)
head (TidyData, 5)
```
                 name                         Population                        Parent Count ParentCount Percent_Parent_Gate_rounded
               <char>                             <char>                        <char> <int>       <int>                       <num>
1: 00_Ctrl.fcs_INF052                           /Scatter                          root 10000       10000                      100.00
2: 00_Ctrl.fcs_INF052                  /Scatter/Singlets                      /Scatter  9986       10000                       99.86
3: 00_Ctrl.fcs_INF052             /Scatter/Singlets/Live             /Scatter/Singlets  9708        9986                       97.22
4: 00_Ctrl.fcs_INF052      /Scatter/Singlets/Live/Tcells        /Scatter/Singlets/Live  9708        9708                      100.00
5: 00_Ctrl.fcs_INF052 /Scatter/Singlets/Live/Tcells/CD4+ /Scatter/Singlets/Live/Tcells  6151        9708                       63.36

# Problem 2:
Instead of removing the files from data folder, I decided to move a couple of them to my folder, inside of the homeworks folder:

```{r}
GS <- flowjo_to_gatingset(ws=ws, name=1, path=Folder)
```

Error:
! Multiple FCS files match sample 00_Ctrl.fcs by filename, event count, and keywords.
Candidates are: 
C:\Users\Macarena\CytometryInR\course\05_GatingSets\data\2025_07_26_AB_02_INF100_00_Ctrl.fcs
C:\Users\Macarena\CytometryInR\course\05_GatingSets\data\2025_07_26_AB_02_INF179_00_Ctrl.fcs
Please move incorrect files out of this directory or its subdirectories.
Show Traceback

This is the same error reported with the 6 files. Let's try with "additional.keys" as GROUPNAME:

```{r}
GS <- flowjo_to_gatingset(ws=ws, name=1, path=Folder, additional.keys="GROUPNAME")
GS
```
FCS not found for sample 00_Ctrl.fcs_INF052 from searching the file extension: .fcs
FCS not found for sample 00_SEB.fcs_INF052 from searching the file extension: .fcs
A GatingSet with 4 samples

It is not an error, but tell us that 2 files are missed.

# Problem 3:
To see the different population cells:
```{r}
 Data <- gs_pop_get_count_fast(GS)
 Data
```

Then, we need to load ggplot2 and ggcyto:

```{r}
library(ggplot2)
library(ggcyto)
```

```{r}
ggcyto(GS[1], subset="root", aes(x="IFNg", y="TNFa")) + geom_hex(bins=125)
```

```{r}
ggcyto(GS[1], subset="Tcells", aes(x="IFNg", y="TNFa")) + geom_hex(bins=125)
```
```{r}
ggcyto(GS[1], subset="CD4+", aes(x="IFNg", y="TNFa")) + geom_hex(bins=130)
```

```{r}
ggcyto(GS[1], subset="CD8+", aes(x="IFNg", y="TNFa")) + geom_hex(bins=105)
```

```{r}
ggcyto(GS[2], subset="root", aes(x="IFNg", y="TNFa")) + geom_hex(bins=110)
```

```{r}
ggcyto(GS[2], subset="Tcells", aes(x="IFNg", y="TNFa")) + geom_hex(bins=110)
```

```{r}
ggcyto(GS[2], subset="CD4+", aes(x="IFNg", y="TNFa")) + geom_hex(bins=120)
```

```{r}
ggcyto(GS[2], subset="CD8+", aes(x="IFNg", y="TNFa")) + geom_hex(bins=105)
```

```{r}
ggcyto(GS[3], subset="root", aes(x="IFNg", y="TNFa")) + geom_hex(bins=100)
```

```{r}
ggcyto(GS[3], subset="Tcells", aes(x="IFNg", y="TNFa")) + geom_hex(bins=110)
```

```{r}
ggcyto(GS[3], subset="CD4+", aes(x="IFNg", y="TNFa")) + geom_hex(bins=110)
```

```{r}
ggcyto(GS[3], subset="CD8+", aes(x="IFNg", y="TNFa")) + geom_hex(bins=90)
```

```{r}
ggcyto(GS[4], subset="root", aes(x="IFNg", y="TNFa")) + geom_hex(bins=115)    
```

```{r}
ggcyto(GS[4], subset="Tcells", aes(x="IFNg", y="TNFa")) + geom_hex(bins=110)
```

```{r}
ggcyto(GS[4], subset="CD4+", aes(x="IFNg", y="TNFa")) + geom_hex(bins=130)
```

```{r}
ggcyto(GS[4], subset="CD8+", aes(x="IFNg", y="TNFa")) + geom_hex(bins=90)
```

```{r}
ggcyto(GS[5], subset="root", aes(x="IFNg", y="TNFa")) + geom_hex(bins=115)    
```

```{r}
ggcyto(GS[5], subset="Tcells", aes(x="IFNg", y="TNFa")) + geom_hex(bins=100)
```

```{r}
ggcyto(GS[5], subset="CD4+", aes(x="IFNg", y="TNFa")) + geom_hex(bins=130)
```

```{r}
ggcyto(GS[5], subset="CD8+", aes(x="IFNg", y="TNFa")) + geom_hex(bins=110)
```

```{r}
ggcyto(GS[6], subset="root", aes(x="IFNg", y="TNFa")) + geom_hex(bins=135)    
```

```{r}
ggcyto(GS[6], subset="Tcells", aes(x="IFNg", y="TNFa")) + geom_hex(bins=120)
```

```{r}
ggcyto(GS[6], subset="CD4+", aes(x="IFNg", y="TNFa")) + geom_hex(bins=130)
```

```{r}
ggcyto(GS[6], subset="CD8+", aes(x="IFNg", y="TNFa")) + geom_hex(bins=120)
```