---
Title: "Gating sets_Homeworks"
Author: "Macarena López-Mayorga"
date: 2026-04-30
format: html
---
# Problem 1
First of all, we are going to indicate the file path of the fcs files:

```{r}
Folder <- file.path("C:/Users/Macarena/CytometryInR/course/05_GatingSets/data")
fcs_files <- list.files(Folder, pattern=".fcs", full.names=TRUE)
fcs_files
```
[1] "C:/Users/Macarena/CytometryInR/course/05_GatingSets/data/2025_07_26_AB_02_INF052_00_Ctrl.fcs"
[2] "C:/Users/Macarena/CytometryInR/course/05_GatingSets/data/2025_07_26_AB_02_INF052_00_SEB.fcs" 
[3] "C:/Users/Macarena/CytometryInR/course/05_GatingSets/data/2025_07_26_AB_02_INF100_00_Ctrl.fcs"
[4] "C:/Users/Macarena/CytometryInR/course/05_GatingSets/data/2025_07_26_AB_02_INF100_00_SEB.fcs" 
[5] "C:/Users/Macarena/CytometryInR/course/05_GatingSets/data/2025_07_26_AB_02_INF179_00_Ctrl.fcs"
[6] "C:/Users/Macarena/CytometryInR/course/05_GatingSets/data/2025_07_26_AB_02_INF179_00_SEB.fcs" 

```{r}
library(flowCore)
```

```{r}
library(lobstr)
library(flowWorkspace)
```

```{r}
flowSet <- read.flowSet(files=fcs_files, truncate_max_range=FALSE, transformation=FALSE)
flowSet
```

```{r}
GatingSet1 <- GatingSet(flowSet)
GatingSet1
```

```{r}
library(CytoML)
```

```{r}
Folder <- file.path("C:/Users/Macarena/CytometryInR/course/05_GatingSets/data")
FlowJoWsp <- list.files(path=Folder, pattern=".wsp", full=TRUE)
FlowJoWsp
```

```{r}
OpenedWsp <- FlowJoWsp[stringr::str_detect(FlowJoWsp, "Opened")]
OpenedWsp
```

```{r}
ws <- open_flowjo_xml(OpenedWsp)
ws
```

```{r}
GS <- flowjo_to_gatingset(ws=ws, name=1, path=Folder, additional.keys="GROUPNAME")
GS
```

```{r}
library(bench)
```

```{r}
Data <- gs_pop_get_count_fast(GS)
head(Data, 5)
```
                 name                         Population                        Parent Count ParentCount
               <char>                             <char>                        <char> <int>       <int>
1: 00_Ctrl.fcs_INF052                           /Scatter                          root 10000       10000
2: 00_Ctrl.fcs_INF052                  /Scatter/Singlets                      /Scatter  9986       10000
3: 00_Ctrl.fcs_INF052             /Scatter/Singlets/Live             /Scatter/Singlets  9708        9986
4: 00_Ctrl.fcs_INF052      /Scatter/Singlets/Live/Tcells        /Scatter/Singlets/Live  9708        9708
5: 00_Ctrl.fcs_INF052 /Scatter/Singlets/Live/Tcells/CD4+ /Scatter/Singlets/Live/Tcells  6151        9708

```{r}
library(dplyr)
```

```{r}
TidyData <- Data |> mutate(Percent_Parent_Gate=(Count/ParentCount)*100)
TidyData
```

```{r}
TidyData <- TidyData |> mutate(Percent_Parent_Gate_rounded=round(Percent_Parent_Gate, 2))
head(TidyData, 5)
```