---
title: week2 homeworks
author: Macarena López-Mayorga
date: 2026-02-26
format: html
---
# Problem 1
First of all, I want to check where we are

```{r}
#| eval: FALSE
getwd()
```
[1] "c:/Users/Macarena/CytometryInR"

So, we are at the top, in the folder "CytometryInR". I'm going to use the file path function to indicate where is the USB unit

```{r}
#| eval: FALSE
FoldersLocation <- file.path("E:")
FoldersLocation
```
[1] "E:"

Now, I want to see the files in the USB unit

```{r}
#| eval: FALSE
list.files(path=FoldersLocation, full.names=FALSE, recursive=FALSE)
```
 [1] "all data microarrays"                         "Back Up Your Files to the Cloud.pdf"          "CABD meeting 01062020"                       
 [4] "CABD_meeting01062020-9.pptx"                  "craneos  Bex3"                                "DE-standard-HOXD13-RBFOX2_29072015.pcrd"     
 [7] "DE-standard-HOXD13-RBFOX2_29072015.pltd"      "Desktop DB"                                   "Desktop DF"                                  
[10] "Entrega_GENext99-18_raw data microarrays EEE" "figuras y pies pdf"                           "FILEID.DAT"                                  
[13] "FINDER.DAT"                                   "fotos Claudia imprimir"                       "Macarena pics backup 28032019"               
[16] "New pics ISH PAE-LIMBS"                       "pics backup 08052019"                         "pics backup 29082018"                        
[19] "presupuestos y vales pedido"                  "SanDiskSecureAccess"                          "SanDiskSecureAccessV3.1_win.exe"             
[22] "sdh"                                          "System Volume Information"                    "TargetFolder"                                
[25] "TEMA_II_ETICA_V2018.ppt"                      "TEMA_II_ETICA_V2019MLM.ppt"                   "Tema_IV_Reconocimiento_Dolor_2018.ppt"       
[28] "Temario_especifico"                           "tesis"                                        "TheVolumeSettingsFolder"     

```{r}
#| eval: FALSE
# To check that, indeed, there is no files in TargetFolder
FilesPresent <- list.files("E:/TargetFolder", full.names=FALSE, recursive=FALSE)
FilesPresent
```
character(0)

So, R is able to "see" the files and folders in the USB unit. Now, I have a folder in my desktop, FCS DATA, with some fcs experiments. I'm going to copy some of them to the TargetFolder, in the USB unit.


```{r}
#| eval: FALSE
# First, I want to name them and see the path and check if everything goes well
MyFCSfiles <- list.files("C:/Users/Macarena/Desktop/FCS DATA/020425 3.0", pattern=".fcs", full.names=TRUE, recursive=FALSE)
MyFCSfiles
```
 [1] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/1 SP BLANCO_Data Source - 1.fcs"                      
 [2] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/10 hBO CD56 PercP-Cy5.5_Data Source - 1.fcs"          
 [3] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/11 hBO LVDV +CD56 PercP-Cy5.5 - 1_Data Source - 1.fcs"
 [4] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/2 SP CONTROL DAPI_Data Source - 1.fcs"                
 [5] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/3 SP CD3 BV785 _Data Source - 1.fcs"                  
 [6] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/4 SP CD172 PercP-eF710_Data Source - 1.fcs"           
 [7] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/5 SP CD3+DAPI_Data Source - 1.fcs"                    
 [8] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/6 SP CD172+DAPI_Data Source - 1.fcs"                  
 [9] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/7 hBO BLANCO_Data Source - 1.fcs"                     
[10] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/8 hBO LIVEDEAD_Data Source - 1.fcs"                   
[11] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/9 hBO CD56 PercP-Cy5.5_Data Source - 1.fcs"   

```{r}
#| eval: FALSE
# Now, I'm going to copy them to the folder in the USB unit
file.copy(from=MyFCSfiles, to="E:/TargetFolder")
```
 [1] TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE

 They are already copied into the TargetFolder, in the USB unit

```{r}
#| eval: FALSE
# to check the files in the TargetFolder
FilesCopied <- list.files("E:/TargetFolder", full.names=FALSE, recursive=FALSE)
FilesCopied
```
 [1] "1 SP BLANCO_Data Source - 1.fcs"                       "10 hBO CD56 PercP-Cy5.5_Data Source - 1.fcs"          
 [3] "11 hBO LVDV +CD56 PercP-Cy5.5 - 1_Data Source - 1.fcs" "2 SP CONTROL DAPI_Data Source - 1.fcs"                
 [5] "3 SP CD3 BV785 _Data Source - 1.fcs"                   "4 SP CD172 PercP-eF710_Data Source - 1.fcs"           
 [7] "5 SP CD3+DAPI_Data Source - 1.fcs"                     "6 SP CD172+DAPI_Data Source - 1.fcs"                  
 [9] "7 hBO BLANCO_Data Source - 1.fcs"                      "8 hBO LIVEDEAD_Data Source - 1.fcs"                   
[11] "9 hBO CD56 PercP-Cy5.5_Data Source - 1.fcs" 

# Problem 2
Let's see all files and folders present in "FCS DATA", in my desktop

```{r}
#| eval: FALSE
AllFiles <- list.files("C:/Users/Macarena/Desktop/FCS DATA", full.names=TRUE, recursive=TRUE)
AllFiles
```
 [1] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/1 SP BLANCO_Data Source - 1.fcs"                                                    
 [2] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/10 hBO CD56 PercP-Cy5.5_Data Source - 1.fcs"                                        
 [3] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/11 hBO LVDV +CD56 PercP-Cy5.5 - 1_Data Source - 1.fcs"                              
 [4] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/2 SP CONTROL DAPI_Data Source - 1.fcs"                                              
 [5] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/3 SP CD3 BV785 _Data Source - 1.fcs"                                                
 [6] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/4 SP CD172 PercP-eF710_Data Source - 1.fcs"                                         
 [7] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/5 SP CD3+DAPI_Data Source - 1.fcs"                                                  
 [8] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/6 SP CD172+DAPI_Data Source - 1.fcs"                                                
 [9] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/7 hBO BLANCO_Data Source - 1.fcs"                                                   
[10] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/8 hBO LIVEDEAD_Data Source - 1.fcs"                                                 
[11] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/9 hBO CD56 PercP-Cy5.5_Data Source - 1.fcs"                                         
[12] "C:/Users/Macarena/Desktop/FCS DATA/020425 3.0/FCS_Floreada.xml"                                                                   
[13] "C:/Users/Macarena/Desktop/FCS DATA/020425 RATON Y EMBRIOIDES 3.1/020425 3.1/1 SP BLANCO_Data Source - 1.fcs"                      
[14] "C:/Users/Macarena/Desktop/FCS DATA/020425 RATON Y EMBRIOIDES 3.1/020425 3.1/10 hBO CD56 PercP-Cy5.5_Data Source - 1.fcs"          
[15] "C:/Users/Macarena/Desktop/FCS DATA/020425 RATON Y EMBRIOIDES 3.1/020425 3.1/11 hBO LVDV +CD56 PercP-Cy5.5 - 1_Data Source - 1.fcs"
[16] "C:/Users/Macarena/Desktop/FCS DATA/020425 RATON Y EMBRIOIDES 3.1/020425 3.1/2 SP CONTROL DAPI_Data Source - 1.fcs"                
[17] "C:/Users/Macarena/Desktop/FCS DATA/020425 RATON Y EMBRIOIDES 3.1/020425 3.1/3 SP CD3 BV785 _Data Source - 1.fcs"                  
[18] "C:/Users/Macarena/Desktop/FCS DATA/020425 RATON Y EMBRIOIDES 3.1/020425 3.1/4 SP CD172 PercP-eF710_Data Source - 1.fcs"           
[19] "C:/Users/Macarena/Desktop/FCS DATA/020425 RATON Y EMBRIOIDES 3.1/020425 3.1/5 SP CD3+DAPI_Data Source - 1.fcs"                    
[20] "C:/Users/Macarena/Desktop/FCS DATA/020425 RATON Y EMBRIOIDES 3.1/020425 3.1/6 SP CD172+DAPI_Data Source - 1.fcs"                  
[21] "C:/Users/Macarena/Desktop/FCS DATA/020425 RATON Y EMBRIOIDES 3.1/020425 3.1/7 hBO BLANCO_Data Source - 1.fcs"                     
[22] "C:/Users/Macarena/Desktop/FCS DATA/020425 RATON Y EMBRIOIDES 3.1/020425 3.1/8 hBO LIVEDEAD_Data Source - 1.fcs"                   
[23] "C:/Users/Macarena/Desktop/FCS DATA/020425 RATON Y EMBRIOIDES 3.1/020425 3.1/9 hBO CD56 PercP-Cy5.5_Data Source - 1.fcs"           
[24] "C:/Users/Macarena/Desktop/FCS DATA/curso citometría en R/FCS_Floreada.xml"                                                        
[25] "C:/Users/Macarena/Desktop/FCS DATA/curso citometría en R/FCS_FlowJO.wsp"      

Searching online I found this code that seems the most easiest one to remember
```{r}
#| eval: FALSE
AllDirectories <- list.dirs("C:/Users/Macarena/Desktop/FCS DATA", recursive = FALSE, full.names = FALSE)
AllDirectories
```
[1] "020425 3.0"                    "020425 RATON Y EMBRIOIDES 3.1" "curso citometría en R"  

With this code, R only show you the folders in "FCS DATA"

# Problem 3
First, I need to know the path of FlowSOM and flowWorkspace

```{r}
#| eval: FALSE
system.file("extdata", package = "FlowSOM")
```
[1] "C:/Users/Macarena/AppData/Local/R/win-library/4.5/FlowSOM/extdata"


```{r}
#| eval: FALSE
system.file("extdata", package = "flowWorkspace")
```
[1] "C:/Users/Macarena/AppData/Local/R/win-library/4.5/flowWorkspace/extdata"


```{r}
#| eval: FALSE
# Let's see what files are in the FlowSOM directory
FlowSOMFiles <- list.files("C:/Users/Macarena/AppData/Local/R/win-library/4.5/FlowSOM/extdata", full.names=TRUE, recursive=FALSE)
FlowSOMFiles
```
[1] "C:/Users/Macarena/AppData/Local/R/win-library/4.5/FlowSOM/extdata/68983.fcs"       
[2] "C:/Users/Macarena/AppData/Local/R/win-library/4.5/FlowSOM/extdata/gating.wsp"      
[3] "C:/Users/Macarena/AppData/Local/R/win-library/4.5/FlowSOM/extdata/gatingResult.csv"
[4] "C:/Users/Macarena/AppData/Local/R/win-library/4.5/FlowSOM/extdata/manualGating.xml"


```{r}
#| eval: FALSE
# Now, I name the path "FlowSOMFiles"
FlowSOMFiles <- file.path ("C:/Users/Macarena/AppData/Local/R/win-library/4.5/FlowSOM/extdata")
FlowSOMFiles
```
```{r}
#| eval: FALSE
# Finally, I just want to see the fcs file
list.files(path=FlowSOMFiles, pattern=".fcs", full.names=TRUE, recursive=TRUE)
```
[1] "C:/Users/Macarena/AppData/Local/R/win-library/4.5/FlowSOM/extdata/68983.fcs"


```{r}
#| eval: FALSE
# I repeat the same process for flowWorkspace
flowWorkspaceFilesList <- list.files("C:/Users/Macarena/AppData/Local/R/win-library/4.5/flowWorkspace/extdata", full.names=TRUE, recursive=TRUE)
flowWorkspaceFilesList
```
[1] "C:/Users/Macarena/AppData/Local/R/win-library/4.5/flowWorkspace/extdata/empty_workspace_9.2.xml"

In this case, it seems that there is no fcs file in the flowWorkspace directory, but I am going to look specifically for fcs files

```{r}
#| eval: FALSE
flowWorkspaceFilesFCS <- file.path("C:/Users/Macarena/AppData/Local/R/win-library/4.5/flowWorkspace/extdata")
flowWorkspaceFilesFCS
```

```{r}
list.files(path=flowWorkspaceFilesFCS, pattern=".fcs", full.names=TRUE, recursive=TRUE)
```
character(0)
So, indeed, there is no fcs files in this directory