---
Title: "week07-Applying Transformations"
Author: "Macarena López-Mayorga"
date: 2026-05-27
format: html
---
First of all, I am going to load in flowWorkspace and ggcyto packages:

```{r}
library(flowWorkspace)
library(ggcyto)
library(dplyr)
library(stringr)
```
Now, I have to indicate the file path and the files that I am going to use (I am going to use just the fcs files)

```{r}
StorageLocation <- file.path("C:/Users/Macarena/CytometryInR/course/07_Transformations/data")
fcs_files <- list.files(StorageLocation, ".fcs", full.names=TRUE)
SFC_files <- fcs_files[stringr::str_detect(fcs_files, "2025")]
```

```{r}
SFC_files
```
[1] "C:/Users/Macarena/CytometryInR/course/07_Transformations/data/2025_07_26_AB_02_INF052_00_Ctrl.fcs"
[2] "C:/Users/Macarena/CytometryInR/course/07_Transformations/data/2025_07_26_AB_02_INF052_00_SEB.fcs" 
[3] "C:/Users/Macarena/CytometryInR/course/07_Transformations/data/2025_07_26_AB_02_INF100_00_Ctrl.fcs"
[4] "C:/Users/Macarena/CytometryInR/course/07_Transformations/data/2025_07_26_AB_02_INF100_00_SEB.fcs" 
[5] "C:/Users/Macarena/CytometryInR/course/07_Transformations/data/2025_07_26_AB_02_INF179_00_Ctrl.fcs"
[6] "C:/Users/Macarena/CytometryInR/course/07_Transformations/data/2025_07_26_AB_02_INF179_00_SEB.fcs" 

Next, I'm going to list the transformer objects
```{r}
SFC_cytoset <- load_cytoset_from_fcs(SFC_files, truncate_max_range=FALSE, transformation=FALSE)
SFC_GatingSet <- GatingSet(SFC_cytoset)
SFC_GatingSet
```
A GatingSet with 6 samples

And I'm going to apply transformers

```{r}
Logicle <- logicle_trans()
str(Logicle)
```
List of 9
 $ name        : chr "logicle"
 $ transform   :function (x)  
 $ inverse     :function (x)  
 $ d_transform : NULL
 $ d_inverse   : NULL
 $ breaks      :function (x)  
 $ minor_breaks:function (b, limits, n)  
 $ format      :function (x)  
 $ domain      : num [1:2] -Inf Inf
 - attr(*, "class")= chr "transform"
 
 ```{r}
 Biexponential <- flowjo_biexp_trans()
 str(Biexponential)
 ```
 List of 9
 $ name        : chr "flowJo_biexp"
 $ transform   :function (x, deriv = 0)  
  ..- attr(*, "type")= chr "biexp"
  ..- attr(*, "parameters")=List of 5
  .. ..$ channelRange: int 4096
  .. ..$ maxValue    : num 262144
  .. ..$ neg         : num 0
  .. ..$ pos         : num 4.5
  .. ..$ widthBasis  : num -10
 $ inverse     :function (x, deriv = 0)  
  ..- attr(*, "type")= chr "biexp"
  ..- attr(*, "parameters")=List of 5
  .. ..$ channelRange: int 4096
  .. ..$ maxValue    : num 262144
  .. ..$ neg         : num 0
  .. ..$ pos         : num 4.5
  .. ..$ widthBasis  : num -10
 $ d_transform : NULL
 $ d_inverse   : NULL
 $ breaks      :function (x)  
 $ minor_breaks:function (b, limits, n)  
 $ format      :function (x)  
 $ domain      : num [1:2] -Inf Inf
 - attr(*, "class")= chr "transform"
 
 ```{r}
 Asinh <- flowjo_fasinh_trans()
 str(Asinh)
 ```
 List of 9
 $ name        : chr "flowJo_fasinh"
 $ transform   :function (x)  
 $ inverse     :function (x)  
 $ d_transform : NULL
 $ d_inverse   : NULL
 $ breaks      :function (x)  
 $ minor_breaks:function (b, limits, n)  
 $ format      :function (x)  
 $ domain      : num [1:2] -Inf Inf
 - attr(*, "class")= chr "transform"
 
 ```{r}
 AsinhGML <- asinhtGml2_trans()
 str(AsinhGML)
 ```
 List of 9
 $ name        : chr "asinhtGml2"
 $ transform   :function (x)  
 $ inverse     :function (x)  
 $ d_transform : NULL
 $ d_inverse   : NULL
 $ breaks      :function (x)  
 $ minor_breaks:function (b, limits, n)  
 $ format      :function (x)  
 $ domain      : num [1:2] -Inf Inf
 - attr(*, "class")= chr "transform"
 
 To be sure that what we have to transform:
 ```{r}
 SFC_Parameters <- colnames(SFC_GatingSet)
 SFC_Parameters
 ```
  [1] "Time"              "SSC-W"             "SSC-H"             "SSC-A"             "FSC-W"             "FSC-H"             "FSC-A"             "SSC-B-W"          
 [9] "SSC-B-H"           "SSC-B-A"           "BUV395-A"          "BUV563-A"          "BUV615-A"          "BUV661-A"          "BUV737-A"          "BUV805-A"         
[17] "Pacific Blue-A"    "BV480-A"           "BV570-A"           "BV605-A"           "BV650-A"           "BV711-A"           "BV750-A"           "BV786-A"          
[25] "Alexa Fluor 488-A" "Spark Blue 550-A"  "Spark Blue 574-A"  "RB613-A"           "RB705-A"           "RB780-A"           "PE-A"              "PE-Dazzle594-A"   
[33] "PE-Cy5-A"          "PE-Fire 700-A"     "PE-Fire 744-A"     "PE-Vio770-A"       "APC-A"             "Alexa Fluor 647-A" "APC-R700-A"        "Zombie NIR-A"     
[41] "APC-Fire 750-A"    "APC-Fire 810-A"    "AF-A"   

```{r}

```