---
title: "Homeworks week 3"
author: "Macarena López-Mayorga"
date: 2026-03-16
format: html
---
# Problem 1
First, I am going to call the file path
```{r}
#| eval: FALSE
UnmixedFile <- file.path("C:/Users/Macarena/CytometryInR/course/03_InsideFCSFile/data/AdditionalFCSFiles/2025_07_26_AB_02_NY068_02_Ctrl.fcs")
UnmixedFile
```

Next, I need to activate flowCore

```{r}
#| eval: FALSE
library(flowCore)
```

Now, let's see the filename argument
```{r}
#| eval= FALSE
read.FCS(filename=UnmixedFile)
```
flowFrame object '2025_07_26_AB_02_NY068_02_Ctrl.fcs'
with 100 cells and 43 observables:
               name      desc     range  minRange    maxRange
$P1            Time        NA    516839         0     506.501
$P2           SSC-W        NA   4194304         0 4194303.000
$P3           SSC-H        NA   4194304         0 4194303.000
$P4           SSC-A        NA   4194304         0 4194303.000
$P5           FSC-W        NA   4194304         0 4194303.000
...             ...       ...       ...       ...         ...
$P39     APC-R700-A    CD107a   4194304      -111     4192506
$P40   Zombie NIR-A Viability   4194304      -111     4192506
$P41 APC-Fire 750-A      CD27   4194304      -111     4192506
$P42 APC-Fire 810-A      CCR7   4194304      -111     4192506
$P43           AF-A        NA   4194304      -111     4194303
472 keywords are stored in the 'description' slot

This is the argument to see the different slots
```{r}
#| eval= FALSE
flowFrame <- read.FCS(filename=UnmixedFile, transformation=FALSE, truncate_max_range=FALSE)
```

Let's see the exprs slot

```{r}
#| eval: FALSE
MFI_Matrix <- flowFrame@exprs
```