---
title: week3_inside an FCS file
author: Macarena López-Mayorga
date: 2026-03-05
format: html
---
We are going to work with some fcs files located in the week3. First, we have to locate them and name them

```{r}
#| eval: FALSE
PathToDataFolder <- file.path("C:/Users/Macarena/CytometryInR/course/03_InsideFCSFile/data")
> PathToDataFolder
```
[1] "C:/Users/Macarena/CytometryInR/course/03_InsideFCSFile/data"

```{r}
#| eval: FALSE
fcs_files <- list.files(PathToDataFolder, full.names=TRUE, pattern=".fcs")
> fcs_files
```
[1] "C:/Users/Macarena/CytometryInR/course/03_InsideFCSFile/data/CellCounts4L_AB_05_ND050_05.fcs"

Next, we are going to use flowCore. We have to call it using library function

```{r}
#| eval: FALSE
library(flowCore)
```

we are going to use the read.fcs function. Let's see the help documentation

```{r}
#|eval: FALSE
?read.FCS
```


```{r}
#| eval: FALSE
firstfile <- fcs_files[1]
firstfile
```

For the function read.FCS, let's use just the first argument for the moment

```{r}
#| eval: FALSE
read.FCS(filename=firstfile)
```
flowFrame object 'CellCounts4L_AB_05-ND050-05.fcs'
with 100 cells and 61 observables:
       name   desc     range  minRange  maxRange
$P1    Time     NA    272140         0    272139
$P2   UV1-A     NA   4194304      -111   4194303
$P3   UV2-A     NA   4194304      -111   4194303
$P4   UV3-A     NA   4194304      -111   4194303
$P5   UV4-A     NA   4194304      -111   4194303
...     ...    ...       ...       ...       ...
$P57   R4-A     NA   4194304      -111   4194303
$P58   R5-A     NA   4194304      -111   4194303
$P59   R6-A     NA   4194304      -111   4194303
$P60   R7-A     NA   4194304      -111   4194303
$P61   R8-A     NA   4194304      -111   4194303
599 keywords are stored in the 'description' slot

R gives back what resembles a table, with the number of cells, the name of files, the name of detectors (column labelled as "name")


```{r}
#| eval: FALSE
flowFrame <- read.FCS(filename=firstfile, transformation = FALSE, truncate_max_range = FALSE)
```
Let's see what kind of file is now


```{r}
#| eval: FALSE
class(flowFrame)
```
[1] "flowFrame"
attr(,"package")
[1] "flowCore"
This is a new type of file named as "S4 object". At the right panel, there is a drop-down in the flowFrame file with 3 slots. The first one, "exprs", has a grid symbol. If we click it, we can see a table with the individual detectors and the data of each cell (until 100). To study this matrix in more detail, we are going to assign a value. Because this is a S4 object, we have to indicate the value specifically to this slot.

# *exprs* slot

```{r}
#| eval: FALSE
MFI_Matrix <- flowFrame@exprs
```

Alternatively, we can use the Bioconductor helper function

```{r}
#| eval: FALSE
MFI_Matrix_Alternate <- exprs(flowFrame)
```

With this last code, we can see a few rows, since all the data will be very complicated to work with

```{r}
#| eval: FALSE
head(MFI_Matrix, 5)
```
       Time      UV1-A       UV2-A     UV3-A      UV4-A      UV5-A      UV6-A      UV7-A     UV8-A       UV9-A     UV10-A    UV11-A    UV12-A
[1,]  38823   37.79983 -184.479996 353.87714 1106.22998 1145.18140 2130.21899 4376.34277 3246.7952 32050.65039  8123.2637 1992.5785 1070.3323
[2,]  39780  234.23021  -98.456429  26.43876   70.65833  -89.93541  -29.14263  -26.34649 -162.6675    17.98848   271.9152  154.8575  163.5411
[3,] 267292 -117.96355   40.732426 473.94574 1177.86975 1516.70935 1985.46130 3658.87671 4140.1724 59792.16406 14013.7969 3427.4324 1668.7588
[4,] 128101  289.87671   -2.723389 -54.11960 -163.71489   32.62989 -134.90411  739.71960  402.9025   427.37534   315.6364 -223.0423  145.7121
[5,] 255221 -104.50541  -71.163338 567.57562  610.23627 1416.61328 2868.16040 4034.58789 3234.6626 40126.46484 10325.0371 1974.0907 1033.8450
         UV13-A    UV14-A    UV15-A    UV16-A  SSC-H      SSC-A      V1-A      V2-A         V3-A      V4-A       V5-A      V6-A       V7-A        V8-A
[1,]  956.43573  290.8685 385.49921 670.97687 657613  750760.12 1171.1390  154.5628 1346.4488525 1706.9260 1923.50940  898.2527 3162.55371  83596.5078
[2,]  -32.81524 -104.9198 103.41382  71.41528  83481   81552.85  266.2424  705.2527  244.3218689  341.0508  381.15939   87.1600  151.68785    119.5544
[3,] 1071.14636  730.1430 214.93053 252.75406 890845 1183519.00 1196.0931 1183.1105 2087.0092773  824.6352 1635.27258 1613.9069 4653.16260 176981.6094
[4,]  127.03777 -207.8978 -55.37944 -45.10131  75103   72457.33  227.9926  556.9189   -0.8137281  205.4732   18.12125  179.6371  -69.50061   -132.4348
[5,]  -21.57245  273.6271 960.16290 341.20633 415791  501690.97  717.2498  929.9780 1358.4512939  788.6506 1208.81006 1156.7040 3118.42627 104951.2578
           V9-A      V10-A        V11-A       V12-A      V13-A       V14-A     V15-A     V16-A   FSC-H      FSC-A SSC-B-H   SSC-B-A      B1-A
[1,] 32506.7617 27161.5137  6236.072754 2220.303223 2023.39966  753.589355  510.9540 228.34962 1055905 1217097.50  716733 815959.06  606.6683
[2,]   -79.7691  -109.1783   -73.991196  114.375542  -11.53453  124.986206 -207.3494 -28.96272   79696   83439.11  104575 103132.83  195.2795
[3,] 69236.4297 57626.8984 13175.838867 4534.874023 3434.38989 1995.172363 1321.8030 615.05560 1092481 1453969.38  757351 982038.31 2010.5110
[4,]   129.2739   231.8918    -5.473321    8.792875  -24.62049   -8.212234 -133.7503 -34.32619   64760   60415.23   67955  66806.13 -146.8936
[5,] 42090.0781 34104.3164  7620.552734 3103.544189 2426.43359  650.836304  290.2892 473.32599 1038362 1184479.00  425296 522873.12 1015.5981
           B2-A       B3-A        B4-A        B5-A        B6-A       B7-A        B8-A       B9-A       B10-A       B11-A     B12-A     B13-A
[1,]  416.98294  4172.4712 192400.0938  93929.9375  54236.3320 19342.6445 10507.24219  9498.1270  4465.50928 1668.048096 2199.9475 1581.7345
[2,]  333.25662   332.1675   -230.9639    196.7810    292.8945  -187.2845    70.90886  -334.3563   188.05545 -663.359619 -163.2331   27.9856
[3,] 2150.21826 10106.9551 437801.5625 212176.1562 124294.3594 45068.3008 24289.59180 22500.1914 10624.99219 4684.497559 3471.7749 2727.3904
[4,]  -34.90987   165.7988    675.0156    136.3076    482.8665   133.0948   432.28091   246.8794   -94.44906    2.905877  106.8489  283.3633
[5,]  639.77527  6034.0200 244022.6094 118871.4609  68616.9688 24067.7949 14182.12793 13019.7100  5577.33984 2753.355957 1279.3009 1423.0276
          B14-A      R1-A        R2-A       R3-A       R4-A       R5-A      R6-A       R7-A       R8-A
[1,] 1487.56860  147.1335  129.867630   35.90353  267.23999   49.79849 -732.7097   42.83144  248.56728
[2,]  205.82298 -142.9224   66.516052  113.63218  -94.41375   98.13978  143.4497 -263.28741  -85.83299
[3,] 2371.95850 -128.3749 -105.482544  726.48547   18.87000   95.47879 -194.4526  -84.08820 -301.46066
[4,]   33.24665  127.5455  122.607941   37.83584  -82.87500 -343.83768   82.3745   60.27896  -94.38461
[5,] 1565.72742 -266.4482   -3.350622 -178.39566 -117.10875 -100.10384 -182.0066  184.36417  186.17207


```{r}
#| eval: FALSE
ColumnNames <- colnames(MFI_Matrix)
ColumnNames
```
 $P1N      $P2N      $P3N      $P4N      $P5N      $P6N      $P7N      $P8N      $P9N     $P10N     $P11N     $P12N     $P13N     $P14N     $P15N 
   "Time"   "UV1-A"   "UV2-A"   "UV3-A"   "UV4-A"   "UV5-A"   "UV6-A"   "UV7-A"   "UV8-A"   "UV9-A"  "UV10-A"  "UV11-A"  "UV12-A"  "UV13-A"  "UV14-A" 
    $P16N     $P17N     $P18N     $P19N     $P20N     $P21N     $P22N     $P23N     $P24N     $P25N     $P26N     $P27N     $P28N     $P29N     $P30N 
 "UV15-A"  "UV16-A"   "SSC-H"   "SSC-A"    "V1-A"    "V2-A"    "V3-A"    "V4-A"    "V5-A"    "V6-A"    "V7-A"    "V8-A"    "V9-A"   "V10-A"   "V11-A" 
    $P31N     $P32N     $P33N     $P34N     $P35N     $P36N     $P37N     $P38N     $P39N     $P40N     $P41N     $P42N     $P43N     $P44N     $P45N 
  "V12-A"   "V13-A"   "V14-A"   "V15-A"   "V16-A"   "FSC-H"   "FSC-A" "SSC-B-H" "SSC-B-A"    "B1-A"    "B2-A"    "B3-A"    "B4-A"    "B5-A"    "B6-A" 
    $P46N     $P47N     $P48N     $P49N     $P50N     $P51N     $P52N     $P53N     $P54N     $P55N     $P56N     $P57N     $P58N     $P59N     $P60N 
   "B7-A"    "B8-A"    "B9-A"   "B10-A"   "B11-A"   "B12-A"   "B13-A"   "B14-A"    "R1-A"    "R2-A"    "R3-A"    "R4-A"    "R5-A"    "R6-A"    "R7-A" 
    $P61N 
   "R8-A" 
   
   Above of each receptor, appears the pattern "$P#N", that remind us to the pattern that appears when we ran read.FCS (see above)
   
   ```{r}
   #| eval: FALSE
   str(ColumnNames)
   ```
   Named chr [1:61] "Time" "UV1-A" "UV2-A" "UV3-A" "UV4-A" "UV5-A" "UV6-A" "UV7-A" "UV8-A" "UV9-A" "UV10-A" "UV11-A" "UV12-A" "UV13-A" "UV14-A" ...
 - attr(*, "names")= chr [1:61] "$P1N" "$P2N" "$P3N" "$P4N" ...

 Here, we see that, each value has a corresponding index name as well. Let's see using class function

 
 ```{r}
 #| eval: FALSE
 class(ColumnNames)
 ```
 [1] "character"

 we just see that all is considered as character, but it doesn't inform us about the index. That's why it is best to use several codes, to obtain as much information as possible.
 We can also remove the names and left just the values:
 
 ```{r}
 #| eval: FALSE
 unname(ColumnNames)
 ```
  [1] "Time"    "UV1-A"   "UV2-A"   "UV3-A"   "UV4-A"   "UV5-A"   "UV6-A"   "UV7-A"   "UV8-A"   "UV9-A"   "UV10-A"  "UV11-A"  "UV12-A"  "UV13-A" 
[15] "UV14-A"  "UV15-A"  "UV16-A"  "SSC-H"   "SSC-A"   "V1-A"    "V2-A"    "V3-A"    "V4-A"    "V5-A"    "V6-A"    "V7-A"    "V8-A"    "V9-A"   
[29] "V10-A"   "V11-A"   "V12-A"   "V13-A"   "V14-A"   "V15-A"   "V16-A"   "FSC-H"   "FSC-A"   "SSC-B-H" "SSC-B-A" "B1-A"    "B2-A"    "B3-A"   
[43] "B4-A"    "B5-A"    "B6-A"    "B7-A"    "B8-A"    "B9-A"    "B10-A"   "B11-A"   "B12-A"   "B13-A"   "B14-A"   "R1-A"    "R2-A"    "R3-A"   
[57] "R4-A"    "R5-A"    "R6-A"    "R7-A"    "R8-A"  

If we look again to the rigth bar, in the drop-down of flowFrame and click in the first slot (exprs) there is another drop-down less user-friendly than the little grid. The columns are represented by []. Inside, in front of the number, it should be a comma (,)


```{r}
#| eval: FALSE
MFI_Matrix[,1]
```
[1]  38823  39780 267292 128101 255221  79210 196643  83855 109315  26128 114423 120001  71831  70551 197021 239994 252611 223012 152780 171822
 [21] 172611 168464 191503 253015  73885  82221 176641 128533   4117 191632 191229  58093 141776 265894  55593 227555 233212 248578  95165 171934
 [41]   1360 251847 195764 147503 118723   1060  90033 253553 268268  74610  23531 150119 226391 201568 179264  79944 196686 252667 117309   3903
 [61]  77690 195142 229873 254472 179943 236618  68193  87154  28541  78622 155664  50115  40866  70753 260118  12033  96149  20740  37461  73998
 [81] 231939 192329  88649 197664  86006 142486 159539 251298 104864 164090 102380 218968 145182 239323 261272 118979  17202 194277 229284 258723

 
 ```{r}
 #| eval: FALSE
 MFI_Matrix[,61]
 ```
   [1]  248.567276  -85.832993 -301.460663  -94.384613  186.172073 -461.407745  843.507080  277.516113 -106.166855  281.633545  195.927261  818.865723
 [13]  734.996460  209.356476  206.442596  279.859894  518.165222   56.947498  285.751007  857.126343  -94.384613 -213.030518   62.585236  138.409653
 [25]  118.012444  328.255768  -61.635056  185.285233  464.384979    5.637739  -66.385956   31.229273 1198.241211  185.475266  873.279419  457.607025
 [37]  -73.353951   37.880539  729.168640  221.772171 -169.512238  348.272888 -338.391022  845.534119   -4.434176  620.024597  610.269409 -193.900208
 [49]  230.830566  -23.754517  607.102112   14.949510  -34.333195 -169.132172  -96.158287  220.631958  125.297165  -15.202891 -126.057304  193.393448
 [61]   90.203819 -277.706146  590.505615  911.096619  -92.230873  347.259369  135.559113  369.430267  -62.015125 -180.597672 -146.517868  810.440796
 [73]  134.038818 -165.268097  727.711731  -88.746880   62.901962  203.275330  436.196289 -242.676147  -40.857769  222.278946 -170.272385  525.513245
 [85]  -41.491222  176.670258  201.501648  175.530045  329.839386  474.140167  -48.142490 -174.833252   46.052090  357.584656  -26.541714  191.493088
 [97]  211.320190  124.790398 -113.324883  343.268616

 If we use a column number that doesn´t exist:
 
 ```{r}
 #| eval: FALSE
 MFI_Matrix[,350]
 ```
 
Error in `MFI_Matrix[, 350]`:
! subscript out of bounds

It gives us an error message.

So, we know how represent the columns [] with a comma inside before a specific number. For rows, it is also used [] but the order is altered: first, we put a specific number and then, a comma.

```{r}
#| eval: FALSE
MFI_Matrix[1,]
```

        Time         UV1-A         UV2-A         UV3-A         UV4-A         UV5-A         UV6-A         UV7-A         UV8-A         UV9-A 
  38823.00000      37.79983    -184.48000     353.87714    1106.22998    1145.18140    2130.21899    4376.34277    3246.79517   32050.65039 
       UV10-A        UV11-A        UV12-A        UV13-A        UV14-A        UV15-A        UV16-A         SSC-H         SSC-A          V1-A 
   8123.26367    1992.57849    1070.33228     956.43573     290.86853     385.49921     670.97687  657613.00000  750760.12500    1171.13904 
         V2-A          V3-A          V4-A          V5-A          V6-A          V7-A          V8-A          V9-A         V10-A         V11-A 
    154.56281    1346.44885    1706.92603    1923.50940     898.25269    3162.55371   83596.50781   32506.76172   27161.51367    6236.07275 
        V12-A         V13-A         V14-A         V15-A         V16-A         FSC-H         FSC-A       SSC-B-H       SSC-B-A          B1-A 
   2220.30322    2023.39966     753.58936     510.95404     228.34962 1055905.00000 1217097.50000  716733.00000  815959.06250     606.66833 
         B2-A          B3-A          B4-A          B5-A          B6-A          B7-A          B8-A          B9-A         B10-A         B11-A 
    416.98294    4172.47119  192400.09375   93929.93750   54236.33203   19342.64453   10507.24219    9498.12695    4465.50928    1668.04810 
        B12-A         B13-A         B14-A          R1-A          R2-A          R3-A          R4-A          R5-A          R6-A          R7-A 
   2199.94751    1581.73450    1487.56860     147.13348     129.86763      35.90353     267.23999      49.79849    -732.70966      42.83144 
         R8-A 
    248.56728 

    We can also ask for a specific value of the table, by putting the index number of the row and the column. For example, if we wanted the MFI value for the UV1-A detector for the first acquired cell (knowing that UV1-A is the 2nd column):

```{r}
#| eval: FALSE
MFI_Matrix[1,2]
```
UV1-A 
37.79983 

# *parameters* slot

This seems to be a more complex slot, with 4 sub-slots in it

## *varMetadata*
It is a table with a column of metadata names, as we can see when we click in the grid. These look reminiscent of what we saw at the top of the read.FCS() column outputs previously (see above)

## *data*
If we click on the grid, we will see the actual content that was displayed.
Let’s try to retrieve the data contained within this slot and save it as it’s own variable/object within our R session. First, we need to open flowFrame object, then use @ to get inside its parameters slot. Since parameters is also a complex object (AnnotatedDataFrame specifically), we will need to use another @ to get inside its data slot:

```{r}
#| eval: FALSE
ParameterData <- flowFrame@parameters@data

head(ParameterData, 10)
```
head(ParameterData, 10)
      name desc   range   minRange maxRange
$P1   Time <NA>  272140    0.00000   272139
$P2  UV1-A <NA> 4194304 -111.00000  4194303
$P3  UV2-A <NA> 4194304 -111.00000  4194303
$P4  UV3-A <NA> 4194304 -111.00000  4194303
$P5  UV4-A <NA> 4194304 -111.00000  4194303
$P6  UV5-A <NA> 4194304 -111.00000  4194303
$P7  UV6-A <NA> 4194304 -111.00000  4194303
$P8  UV7-A <NA> 4194304  -26.34649  4194303
$P9  UV8-A <NA> 4194304 -111.00000  4194303
$P10 UV9-A <NA> 4194304    0.00000  4194303

We can do the same using the Bioconductor helper function

```{r}
#| eval: FALSE
ParameterData_Alternate <- parameters(flowFrame)@data
head(ParameterData_Alternate, 10)
```
      name desc   range   minRange maxRange
$P1   Time <NA>  272140    0.00000   272139
$P2  UV1-A <NA> 4194304 -111.00000  4194303
$P3  UV2-A <NA> 4194304 -111.00000  4194303
$P4  UV3-A <NA> 4194304 -111.00000  4194303
$P5  UV4-A <NA> 4194304 -111.00000  4194303
$P6  UV5-A <NA> 4194304 -111.00000  4194303
$P7  UV6-A <NA> 4194304 -111.00000  4194303
$P8  UV7-A <NA> 4194304  -26.34649  4194303
$P9  UV8-A <NA> 4194304 -111.00000  4194303
$P10 UV9-A <NA> 4194304    0.00000  4194303

Now, we are going to run the str function

```{r}
#| eval: FALSE
str(ParameterData)
```
'data.frame':	61 obs. of  5 variables:
 $ name    : 'AsIs' Named chr  "Time" "UV1-A" "UV2-A" "UV3-A" ...
  ..- attr(*, "names")= chr [1:61] "$P1N" "$P2N" "$P3N" "$P4N" ...
 $ desc    : 'AsIs' Named chr  NA NA NA NA ...
  ..- attr(*, "names")= chr [1:61] NA NA NA NA ...
 $ range   : num  272140 4194304 4194304 4194304 4194304 ...
 $ minRange: num  0 -111 -111 -111 -111 ...
 $ maxRange: num  272139 4194303 4194303 4194303 4194303 ...

 We can see this class of object is a “data.frame”. This is one of the more common object types in R. Each of the columns appears to be designated by a $ followed by the column name, and then type of column (numeric, character, etc).

 If we are trying to see these columns in R, we notice that data.frame is not like the previous S4 class objets we interacted with, as the @ symbol after doesn’t bring up any suggestions

 ```{r}
 #| eval: FALSE
 ParameterData@
 ```
 
Error:
! Can't parse incomplete input

By contrast, adding the $ we see that R suggests us different column names (name, range, minRange, etc). Similar to what we saw with a matrix, we can subset a data.frame based on the column or row index using square brackets [].

```{r}
#| eval: FALSE
ParameterData[,1]
```
   $P1N      $P2N      $P3N      $P4N      $P5N      $P6N      $P7N      $P8N      $P9N     $P10N     $P11N     $P12N     $P13N 
   "Time"   "UV1-A"   "UV2-A"   "UV3-A"   "UV4-A"   "UV5-A"   "UV6-A"   "UV7-A"   "UV8-A"   "UV9-A"  "UV10-A"  "UV11-A"  "UV12-A" 
    $P14N     $P15N     $P16N     $P17N     $P18N     $P19N     $P20N     $P21N     $P22N     $P23N     $P24N     $P25N     $P26N 
 "UV13-A"  "UV14-A"  "UV15-A"  "UV16-A"   "SSC-H"   "SSC-A"    "V1-A"    "V2-A"    "V3-A"    "V4-A"    "V5-A"    "V6-A"    "V7-A" 
    $P27N     $P28N     $P29N     $P30N     $P31N     $P32N     $P33N     $P34N     $P35N     $P36N     $P37N     $P38N     $P39N 
   "V8-A"    "V9-A"   "V10-A"   "V11-A"   "V12-A"   "V13-A"   "V14-A"   "V15-A"   "V16-A"   "FSC-H"   "FSC-A" "SSC-B-H" "SSC-B-A" 
    $P40N     $P41N     $P42N     $P43N     $P44N     $P45N     $P46N     $P47N     $P48N     $P49N     $P50N     $P51N     $P52N 
   "B1-A"    "B2-A"    "B3-A"    "B4-A"    "B5-A"    "B6-A"    "B7-A"    "B8-A"    "B9-A"   "B10-A"   "B11-A"   "B12-A"   "B13-A" 
    $P53N     $P54N     $P55N     $P56N     $P57N     $P58N     $P59N     $P60N     $P61N 
  "B14-A"    "R1-A"    "R2-A"    "R3-A"    "R4-A"    "R5-A"    "R6-A"    "R7-A"    "R8-A" 

  The individual detectors or fluorophore appear under “name”. For now, based on what we know, the $P# appears to be some sort of name being used as an internal consistent reference to the respective.

  ## *dimLabels*
  Let's see the parameters for dimLabels
  
  ```{r}
  #| eval: FALSE
  flowFrame@parameters@dimLabels
  ```
[1] "rowNames"    "columnNames"

## *classVersion*
We are going to continue with the last slot

```{r}
#| eval: FALSE
flowFrame@parameters@.__classVersion__
```
AnnotatedDataFrame 
           "1.1.0" 
# *Description*
When we drop-down this slot, we see a very long list is opened. To make easier the managment of the list, it could be better to retrieve some of the data via the console. To retrieve the list itself, we would need to access the description slot of the flowFrame object. Since it is a slot, we will need to use the @ accessor.

```{r}
#| eval: FALSE
DescriptionList <- flowFrame@description
DescriptionList
```       
The returned list is a little too large to reasonably explore. We can attempt to subset using the head() function as shown below:

```{r}
#| eval: FALSE
head(DescriptionList, 5)
```
$`$BEGINANALYSIS`
[1] "0"

$`$BEGINDATA`
[1] "33312"

$`$BEGINSTEXT`
[1] "0"

$`$BTIM`
[1] "13:55:29.85"

$`$BYTEORD`
[1] "4,3,2,1"
Now, we just see these 5 elements.

We can also use this other code:

```{r}
#| eval: FALSE
DescriptionList[1:10]
```

And a Bioconductor helper keyword function:

```{r}
#| eval: FALSE
DescriptionList_Alternate <- keyword(flowFrame)
DescriptionList_Alternate
```
If we run the class() function, we can see that DescriptionList is an actual “list”.

```{r}
#| eval: FALSE
class(DescriptionList)
```
[1] "list"

